Rule Synthesis for Rule-based Intrusion Detectors

### 2020-06-02 paper/paper.tex
minor changesRQ1-RQ3

\subsection{Step 2: Rules Creation}
\label{sec:minimization}

@@

RQ2 evaluates the importance of each proposed heuristic function. To
answer that question, we evaluated the use of \tname{} without each of
these functions with the goal of checking how much each of them
contributes to the overall performance of the technique. Recall that
the score a rule $r$ is given by the weighted average of the values
given by each heuristic function, \ie{} the score of $r$ is obtained
with the function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the
weight of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on
$r$, and $N$ is the number of heuristics considered---four, in our
case. To assess the value of each heuristic function, we assigned its
weight to zero---effectively, disabling the function---and compared
the result with \tname\ using default weights, obtained with the
optimization procedure described on
Section~\ref{sec:weight-optimization}.

Table~\ref{table:weights} shows results. Column ``Name'' shows the
name of the attack and column ``base'' shows the result of
\tname\ with all heuristic functions enabled, as in column ``Rank
(step 3)'' from Table~\ref{table:results}. The remaining columns show
results obtained by running the same experiment, as on
Section~\ref{sec:answer-rqone}, but suppressing one heuristic
function. The number corresponds to the order the heuristic function
is introduced on Section~\ref{sec:ranking}. The symbol of an up arrow
($\uparrow$) indicates that the performance of \tname{} on that
configuration is worse when compared to \tname's default
configuration. A down arrow ($\downarrow$), appearing highlighted in
the table, indicates the contrary.

\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{8pt}
  %\setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    %%UDP Flood & - & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2665$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table*}[t!]
  \footnotesize
  \setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:comparison}Alternative techniques.}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lcccccc}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Output} &    
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Evaluated Protocols} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Avail.?} \\
    \midrule
    Vollmer\cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
%    \midrule
    Fallahi\cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
%    \midrule
    Kao\cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
%    \midrule
    Guruprasad\cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
    ICMP, TCP & FP & No\\
%    \midrule
    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
    \bottomrule
    \end{tabular}
    \vspace{-3ex}
\end{table*}

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Creations step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule at the rank for an increasing number
\begin{wrapfigure}[10]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-1ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[11]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-1ex}
  \label{fig:impact-negative}
\end{wrapfigure}

>>>>

\subsection{Step 2: Rules Generation}
\label{sec:minimization}

@@

RQ2 assesses the importance of each heuristic function. To that end,
we compared \tname{}, with all functions enabled, against
\tname\ without each one of those functions. Recall that the score a
rule $r$ is given by the weighted average of the values given by each
heuristic function, \ie{}, the score of $r$ is obtained with the
function $f(r)=\sum_{i}^{N} w_i*f_i(r)/N$, where $w_i$ is the weight
of heuristic $i$, $f_i(r)$ is the score of heuristic $i$ on $r$, and
$N$ is the number of heuristics considered. To assess the value of
each heuristic function, we assigned its weight to zero---effectively,
disabling the function---and compared the result with \tname\ using
default weights, obtained with the optimization procedure described on
Section~\ref{sec:weight-optimization}.

Table~\ref{table:weights} shows results. Column ``Name'' shows the
name of the attack and column ``base'' shows the result of
\tname\ with all heuristic functions enabled, as in column ``Rank
(step 3)'' from Table~\ref{table:results}. The remaining columns show
results obtained by running the same experiment, as on
Section~\ref{sec:answer-rqone}, but suppressing one heuristic
function. The number corresponds to the order the heuristic function
is introduced on Section~\ref{sec:ranking}. The symbol of an up arrow
($\uparrow$) indicates that the performance of \tname{} on that
configuration is worse when compared to ``base'' (as the ranking
increased). A down arrow ($\downarrow$), appearing highlighted in the
table, indicates the contrary.

\begin{table}[t!]
  \footnotesize
  \setlength{\tabcolsep}{8pt}
  %\setlength{\tabcolsep}{5pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:weights}Contribution of heuristic functions.}
  \vspace{-2ex}
  \centering
  \begin{tabular}{lrrrrrr}
    \toprule
    Name &
    base &
    -f1 &
    -f2 &    
    -f3 &
    -f4 \\
    \midrule
    Black Nurse & 2 & 7$\uparrow$ & 2 & 2 & 2 \\    
    Ping Scan & 2 &  8$\uparrow$& 2 & 2 & 2 \\
    SYN Flood & 1 & 1 & 1 & 1 & 1 \\
    2100384 & 2 & 18$\uparrow$& 2 & 2 & 2 \\
    2100399 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    2100401 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100402 & 1 & 12$\uparrow$ & 1 & 1 & 1 \\
    2100408 & 2 & 9$\uparrow$ & 2 & 2 & 2 \\
    2100485 & 1 & 2$\uparrow$ & 1 & 1 & 1 \\
    2100486 & 1 & 3$\uparrow$ & 1 & 1 & 1 \\
    %%UDP Flood & - & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & - & - & - & - & - & -\\
    ColdFusion & 2 & 2 & 2 & 7$\uparrow$ & 5$\uparrow$ \\
    iisadmin access & 1  & 1 & 1 & 1 & 14$\uparrow$\\        
    JBoss JMX & 36 & 36 & 14\cellcolor{lightgray}$\downarrow$ & 34\cellcolor{lightgray}$\downarrow$ & 269$\uparrow$ \\
    scripttag & 1 & 1 & 3$\uparrow$ & 1 & 15$\uparrow$ \\
    WeBid LFI & 29 & 29 & 72$\uparrow$ & 39$\uparrow$ & 41$\uparrow$ \\    
    %%Wordpress Login & 8 & - & - & - & - \\
    .htaccess access & 1  & 1 & 1 & 1 & 11$\uparrow$\\
    .idq access & 2  & 2 & 2 & 1\cellcolor{lightgray}$\downarrow$ & 15$\uparrow$ \\
    /system32/ in URI & 1  & 1 & 4$\uparrow$ & 1 & 15$\uparrow$\\
    Basebuilder LFI & 55  & 55 & 320$\uparrow$ & 112$\uparrow$ & 58$\uparrow$\\
      Tomcat Snoop Access & 6 & 6 & 3$\cellcolor{lightgray}\downarrow$ & 6 & 2,665$\uparrow$\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table*}[t!]
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \renewcommand{\arraystretch}{-0.01}
  \caption{\label{table:comparison}Alternative techniques.}
  \vspace{-2ex}
  \centering
    \begin{tabular}{lcccccc}
    \toprule
    \multicolumn{1}{c}{Paper} &
    \multicolumn{1}{c}{Input} &
    \multicolumn{1}{c}{Output} &    
    \multicolumn{1}{c}{Dataset} &
    \multicolumn{1}{c}{Evaluated Protocols} &
    \multicolumn{1}{c}{Evaluation} &
    \multicolumn{1}{c}{Avail.?} \\
    \midrule
    Vollmer\cite{vollmer-etal-cics2011} & Malign Traffic & List  & Manually crafted & ICMP & TP, FP & No\\
%    \midrule
    Fallahi\cite{fallahi} & Mixed Traffic & One Rule  & \cite{iscx} & HTTP, SSH, TCP & TP, FP, Precision, Recall, F1 & No\\    
%    \midrule
    Kao\cite{Kao2015AutomaticNR} & Malign/Benign Traffic, Rule Dataset &
    One Rule & Manually crafted, \cite{emerging-threats-open} & HTTP & Generating time, FP & No \\
%    \midrule
    Guruprasad\cite{guruprasad}& Malign Traffic & List  & \cite{iscx} \cite{darpa1999} &
    ICMP, TCP & FP & No\\
%    \midrule
    Syrius & Malign/Benign Traffic, Rule Dataset & Ranking & Manually crafted, \cite{emerging-threats-open}, \cite{iscx} & HTTP, TCP, ICMP & Rule List & Yes \\
    \bottomrule
    \end{tabular}
    \vspace{-3ex}
\end{table*}

@@

This section evaluates sensitivity of \tname{} to the amount of
examples provided on input. Recall that \tname\ uses one example
attack to generate the seed rule (at the Reverse Engineering step) and
the remaining attacks to evaluate soundness of the rule (at the Rules
Generation step). It is expected that the number and variety of attacks
play a role in the performance of the technique.  We answered this
question in two parts. First, we analyzed the impact of the amount of
positive examples in results. For that, we selected two cases with the
highest number of examples---idqaccess and scripttag---and measured
the position of the golden rule at the rank for an increasing number
\begin{wrapfigure}[9]{r}{0.28\textwidth}  
  \vspace{-2ex}
  \centering
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/idqaccess/idqaccess.pdf}
    \vspace{-3ex}
    \caption{idqaccess}
  \end{subfigure}%
  \begin{subfigure}{.15\textwidth}
    \centering
    \includegraphics[scale=0.35,trim=15 0 130 0,clip]{R/scripttag/scripttag.pdf}
    \vspace{-3ex}        
    \caption{scripttag}
  \end{subfigure}%
    \vspace{-1ex}          
  \caption{\label{fig:impact-number-attacks}Positive examples.}
  \vspace{-2ex}
\end{wrapfigure}
of examples provided to \tname. For each number of variants, we
sampled five sets of that size to provide as input to the technique.
Figure~\ref{fig:impact-number-attacks} shows results for each
attack we selected. The y-axis of each plot denotes the rank of the
golden rule and x-axis denotes the number of attack examples used to
synthesize rules. Each point in the x-axis is associated with the
boxplot of the distribution for the corresponding number of
variants. Results indicate that the number of examples is important,
the choice of examples is important, and a small number of examples
suffices to reach the best results.

\begin{wrapfigure}[8]{r}{0.28\textwidth}  
  \vspace{-2.5ex}
  \centering
  %% \begin{subfigure}{.2\textwidth}
  %%   \centering
  %%   \includegraphics[scale=0.35,trim=10 0 130 0,clip]{R/coldfusion/coldfusion.pdf}
  %%   \vspace{-1ex}        
  %%   \caption{coldfusion}
  %% \end{subfigure}%
  %% \hspace{-8ex}
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/inc/inc.pdf}
    \vspace{-3ex}    
    \caption{inc}    
  \end{subfigure}%
  \begin{subfigure}{.14\textwidth}
    \centering
    \includegraphics[scale=0.3,trim=15 0 130 0,clip]{R/jsp/jsp.pdf}
    \vspace{-3ex}    
    \caption{jsp}    
  \end{subfigure}%
  \vspace{-1ex}
  \caption{Negative examples.}
  \vspace{-2ex}
  \label{fig:impact-negative}
\end{wrapfigure}
